TE_bilingual_vocabulary_model: Prepare the keepers_data for analysis

### 2021-12-06 01_data_wordbank.Rmd
updated paper

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

@@

In our study, we define developmental level of a child as the number of potentially learnable words. The number of learnable words is determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages. So we need to retrieve the percentile information from Wordbank (Frank et al., 2016).
```{r load_wordbankr}
library(wordbankr)
```

@@

## English CDI-Words and sentences: Vocabulary size at 90th percentile
```{r eng_percentile_info}
english_ws_percentile <-  
  fit_vocab_quantiles(get_administration_data("English (American)", "WS"), 
                      production,
                      quantile = "standard") %>%
  # keep only the 90% percentile
  filter(quantile == 0.90) %>%
  rename(EngWS_90percentile = production)
```

## French CDI-Mots et énoncés: Vocabulary size at 90th percentile
```{r fre_percentile_info}
french_ws_percentile <-  
  fit_vocab_quantiles(get_administration_data("French (Quebecois)", "WS"), 
                      production,
                      quantile = "standard") %>%
  # keep only the 90% percentile
  filter(quantile == 0.90) %>%
  rename(FrWS_90percentile = production)
```

# Merge the two percentile dataframe
Then we merge the two percentile dataframe by age.
```{r merge_percentile}
ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
                       by = "age") %>%
  # compute estimate total vocab by adding the two languages
  mutate(Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
  # keep only necessary columns
  select(age, EngWS_90percentile, FrWS_90percentile, Average_90percentile) %>%
  # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
  rename(age_months_percentile = age)
```

>>>>

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)

@@

In our study, we define developmental level of a child as the number of potentially learnable words. The number of learnable words is determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages. So we need to retrieve the percentile information from Wordbank (Frank et al., 2016).
```{r load_wordbankr, cache = TRUE}
library(wordbankr)
```

@@

## English CDI-Words and sentences: Vocabulary size at 90th percentile
```{r eng_percentile_info, cache = TRUE}
english_ws_percentile <-  
  fit_vocab_quantiles(get_administration_data("English (American)", "WS"), 
                      production,
                      quantile = "standard") %>%
  pivot_wider(names_from = quantile, values_from = production) %>%
  rename(EngWS_90percentile = "0.9",
         EngWS_75percentile = "0.75",
         EngWS_50percentile = "0.5",
         EngWS_25percentile = "0.25",
         EngWS_10percentile = "0.1")

```

## French CDI-Mots et énoncés: Vocabulary size at 90th percentile
```{r fre_percentile_info, cache = TRUE}
french_ws_percentile <-  
  fit_vocab_quantiles(get_administration_data("French (Quebecois)", "WS"), 
                      production,
                      quantile = "standard") %>%
  pivot_wider(names_from = quantile, values_from = production) %>%
  rename(FrWS_90percentile = "0.9",
         FrWS_75percentile = "0.75",
         FrWS_50percentile = "0.5",
         FrWS_25percentile = "0.25",
         FrWS_10percentile = "0.1")

```

# Merge the two percentile dataframe
Then we merge the two percentile dataframe by age.
```{r merge_percentile, cache = TRUE}
ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
                       by = "age") %>%
  # compute estimate total vocab by adding the two languages
  mutate(Average_10percentile = (EngWS_10percentile + FrWS_10percentile)/2,
         Average_25percentile = (EngWS_25percentile + FrWS_25percentile)/2,
         Average_50percentile = (EngWS_50percentile + FrWS_50percentile)/2,
         Average_75percentile = (EngWS_75percentile + FrWS_75percentile)/2,
         Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
  # keep only necessary columns
  select(-c(language.x, form.x, language.y, form.y)) %>%
  # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
  rename(age_months_percentile = age)

```

### 2021-12-06 02_clean_data.Rmd
updated paper

```{r load_data_raw}
# raw archival data from Gonzalez-Barrero, Schott, & Byers-Heinlein’s (2020)
cdi_ws = readRDS(here::here("data_raw","output_analyzeSent.Rdata"))
```

@@

```{r}
# change which infants changed lang_group & create a "exclude.language_group_change" criterion for subsequent step
exclude.lang_group_change <- cdi_ws %>%
  filter(lang_group !="other") %>%
  group_by(baby_id) %>% 
  distinct(ID_testdate, .keep_all = T) %>%
  mutate(visit_num = 1:n()) %>%  
  mutate(changed_language_group_status = if_else(visit_num == 1, NA, 
                                                 lang_group != lag(lang_group))) %>%
  # uncomment the next line to see all visits for babies that have one visit excluded
  # filter(any(changed_language_group_status == TRUE) ) %>% View()
  filter(changed_language_group_status == TRUE) %>% 
  pull(baby_id)

@@

# Check the keepers_data
## How many distinct infants and total number of data points are there?
```{r}
keepers_ws_TE %>% summarize(N_visits = n_distinct(ID_testdate),
                            N_babies = n_distinct(baby_id))
```

## Are there any NAs in the vocabulary measures? (should all return 0)
### Vocabulary score in total vocabulary in English/French, total word vocabulary, and total concept vocabulary
```{r}
keepers_ws_TE %>% filter(is.na(vocab_score)) %>% 
  select(ID_testdate, vocab_type, vocab_score) %>%
  nrow()
```

### Unique words in English
```{r}
keepers_ws_TE %>% filter(is.na(eng_unique_words)) %>% 
  select(ID_testdate) %>% 
  nrow()
```

### Unique words in French
```{r}
keepers_ws_TE %>% filter(is.na(fre_unique_words)) %>% 
  select(ID_testdate) %>%
  nrow()
```

### Number of translation equivalents
```{r}
keepers_ws_TE %>% filter(is.na(number_of_te)) %>% 
  select(ID_testdate) %>%
  nrow()
```

# Save the keepers_data
The keepers_data is saved in the folder "data_keepers"
```{r}
write.csv(keepers_ws_TE, "data_keepers/keepers_ws_TE.csv", row.names = F)
```

>>>>

```{r load_data_raw, cache = TRUE}
# raw archival data from Gonzalez-Barrero, Schott, & Byers-Heinlein’s (2020)
cdi_ws = readRDS(here::here("data_raw","output_analyzeSent.Rdata"))
```

@@

```{r cache = TRUE}
# change which infants changed lang_group & create a "exclude.language_group_change" criterion for subsequent step
exclude.lang_group_change <- cdi_ws %>%
  filter(lang_group !="other") %>%
  group_by(baby_id) %>% 
  distinct(ID_testdate, .keep_all = T) %>%
  mutate(visit_num = 1:n()) %>%  
  mutate(changed_language_group_status = if_else(visit_num == 1, NA, 
                                                 lang_group != lag(lang_group))) %>%
  # uncomment the next line to see all visits for babies that have one visit excluded
  # filter(any(changed_language_group_status == TRUE) ) %>% View()
  filter(changed_language_group_status == TRUE) %>% 
  pull(baby_id)

@@

# Check the keepers_data
## How many distinct infants and total number of data points are there?
```{r cache = TRUE}
keepers_ws_TE %>% summarize(N_visits = n_distinct(ID_testdate),
                            N_babies = n_distinct(baby_id))
```

## Are there any NAs in the vocabulary measures? (should all return 0)
### Vocabulary score in total vocabulary in English/French, total word vocabulary, and total concept vocabulary
```{r cache = TRUE}
keepers_ws_TE %>% filter(is.na(vocab_score)) %>% 
  select(ID_testdate, vocab_type, vocab_score) %>%
  nrow()
```

### Unique words in English
```{r cache = TRUE}
keepers_ws_TE %>% filter(is.na(eng_unique_words)) %>% 
  select(ID_testdate) %>% 
  nrow()
```

### Unique words in French
```{r cache = TRUE}
keepers_ws_TE %>% filter(is.na(fre_unique_words)) %>% 
  select(ID_testdate) %>%
  nrow()
```

### Number of translation equivalents
```{r cache = TRUE}
keepers_ws_TE %>% filter(is.na(number_of_te)) %>% 
  select(ID_testdate) %>%
  nrow()
```

# Save the keepers_data
The keepers_data is saved in the folder "data_keepers"
```{r cache = TRUE}
write.csv(keepers_ws_TE, "data_keepers/keepers_ws_TE.csv", row.names = F)
```

### 2021-12-06 03_prepare_data.Rmd
updated paper

```{r load_data_raw}
keepers_ws_TE <- read.csv(here::here("data_keepers/keepers_ws_TE.csv"))
```

# Pivot data from long to wide
```{r}
keepers_ws_TE_final <- keepers_ws_TE %>%
  # remove irrelevant variables
  select(-c(vocab_type_dom, vocab_cross_lang, n_administrations)) %>% 
  # change data from long to wide format by spreading "vocab_type"
  pivot_wider(names_from = vocab_type, values_from = vocab_score)
```

@@

```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # define lang_dom_vocab
  mutate(lang_dom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  # make lang_dom_vocab a factor with 2 levels
  mutate(lang_dom_vocab = as.factor(lang_dom_vocab)) 
```

@@

```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  mutate(
    # define total number of words in the DOMINANT language
    total_words_dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                            TRUE ~ total_words_fre),
    # define total number of words in the NON-DOMINANT language
    total_words_nondom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                               TRUE ~ total_words_eng),
    # define total number of singlets in the DOMINANT language
    total_singlet_dom = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                  TRUE ~ fre_unique_words),
    # define total number of singlets in the NON-DOMINANT language
    total_singlet_nondom = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                     TRUE ~ eng_unique_words),
    # compute total number of singlets across both languages
    singlet_vocab = total_singlet_dom + total_singlet_nondom)
```

@@

```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # calculate Vocabulary Balance balance_vocab
  mutate(balance_vocab = total_words_nondom/(total_words_dom+total_words_nondom)) %>%
  # calculate Input Balance balance_input
  mutate(
    # define input dominance (with reference to the lang_dom_vocab defined by vocab size)
    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_eng,
                              TRUE ~ lang_exp_fre),
    lang_nondom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_fre,
                                 TRUE ~ lang_exp_eng),
    # calculate the balance score
    balance_input = lang_nondom_input/(lang_dom_input+lang_nondom_input))
```

@@

```{r}
# read in the percentile data
ws_percentile <- read.csv(here::here("data_keepers/ws_percentile.csv"))

@@

# Save the final keepers_data
The final keepers_data is saved in the folder "data_keepers"
```{r}
write.csv(keepers_ws_TE_final, "data_keepers/keepers_ws_TE_final.csv", row.names = F)
```

>>>>

```{r load_data_raw, cache = TRUE}
keepers_ws_TE <- read.csv(here::here("data_keepers/keepers_ws_TE.csv"))
```

# Pivot data from long to wide
```{r cache = TRUE}
keepers_ws_TE_final <- keepers_ws_TE %>%
  # remove irrelevant variables
  select(-c(vocab_type_dom, vocab_cross_lang, n_administrations)) %>% 
  # change data from long to wide format by spreading "vocab_type"
  pivot_wider(names_from = vocab_type, values_from = vocab_score)
```

@@

```{r cache = TRUE}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # define lang_dom_vocab
  mutate(lang_dom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  # make lang_dom_vocab a factor with 2 levels
  mutate(lang_dom_vocab = as.factor(lang_dom_vocab)) 
```

@@

```{r cache = TRUE}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  mutate(
    # define total number of words in the DOMINANT language
    total_words_dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                            TRUE ~ total_words_fre),
    # define total number of words in the NON-DOMINANT language
    total_words_nondom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                               TRUE ~ total_words_eng),
    # define total number of singlets in the DOMINANT language
    total_singlet_dom = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                  TRUE ~ fre_unique_words),
    # define total number of singlets in the NON-DOMINANT language
    total_singlet_nondom = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                     TRUE ~ eng_unique_words),
    # compute total number of singlets across both languages
    singlet_vocab = total_singlet_dom + total_singlet_nondom)
```

@@

```{r cache = TRUE}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # calculate Vocabulary Balance balance_vocab
  mutate(balance_vocab = total_words_nondom/(total_words_dom+total_words_nondom)) %>%
  # calculate Input Balance balance_input
  mutate(
    # define input dominance (with reference to the lang_dom_vocab defined by vocab size)
    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_eng,
                              TRUE ~ lang_exp_fre),
    lang_nondom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_fre,
                                 TRUE ~ lang_exp_eng),
    # calculate the balance score
    balance_input = lang_nondom_input/(lang_dom_input+lang_nondom_input))
```

@@

```{r cache = TRUE}
# read in the percentile data
ws_percentile <- read.csv(here::here("data_keepers/ws_percentile.csv"))

@@

# Save the final keepers_data
The final keepers_data is saved in the folder "data_keepers"
```{r cache = TRUE}
write.csv(keepers_ws_TE_final, "data_keepers/keepers_ws_TE_final.csv", row.names = F)
```
